% TODO L01 INTRODUCTION

\ifuniversity{tubs}{\date{October 15, 2024}}

\author{Thomas Thüm}
\lecture{Introduction}{introduction}
% goal of this lecture?
% learn what software is and distinguish term from program, software product, app, ...
% understand the importance/role of software
% understand what software engineering is good for

\inputuniversity{content/01-prelude}

\section{The Impact of Software}
\input{content/01a-software}
\lessonslearned{
	\item What is software?
	\item What is the difference between program, software product, professional software, desktop/web/mobile app?
	\item What are characteristics and properties of software?
	\item What is the impact of software?
	\item Next: What can go possibly wrong with software?
}{
	\item \sommerville\mychapter{1.1}\mypages{19--28} % TODO check
}{
	\begin{enumerate}
		\item<+-> Form groups of 2-3 students
		\item<+-> Introduce yourselves to each other (5 min)
		\item<+-> Share your (recent) experiences with software (5 min)
		\item<+-> Survey: What is your experience with software?\\
		\slideonly{~
			\item<+->[A] Warst du jemals von einer Softwarepanne betroffen?
			\item<+->[B] Hast du dich im letzten Semester über Softwarefehler geärgert?
			\item<+->[C] Hast du selbst Programmiererfahrung?
			\item<+->[D] Hast du bereits Software getestet?
			\item<+->[E] Hast du schon mal Programmierfehler verursacht?
		}
			%\fancyqr{https://studip.tu-braunschweig.de/dispatch.php/questionnaire/answer/dc864612bd163af167a09edabdf058f3?cid=c37620ed3e63513af37fbe31060fcf69&range_type=course&range_id=c37620ed3e63513af37fbe31060fcf69}
	\end{enumerate}
}

\section{Software and Its Engineering}
% Why is Engineering Good Software so Hard?
% How to Develop Good Software?
% Why is Software Production so hard?
\input{content/01b-challenges}
\lessonslearned{
	\item What is software engineering?
	\item Which trade-off is crucial to software engineering?
	\item Next: What is expected from you during this term?
}{
	\item \sommerville\mychapter{1.1}\mypages{19--28} % TODO check
}{
	\begin{enumerate}
		\item<+-> Form groups of 2-3 students
		\item<+-> Discuss: What will you do for a living in 10 years? What will be your connection to software? (5 min)
		\item<+-> Survey: What is your connection to software in 10 years?
		\slideonly{~
			\item<+->[A] Wirst du in 10 Jahren Software entwickeln?
			\item<+->[B] Wirst du in 10 Jahren Software testen?
			\item<+->[C] Wirst du in 10 Jahren Software beauftragen?
			\item<+->[D] Wirst du in 10 Jahren Softwareanforderungen ermitteln?
			\item<+->[E] Wirst du in 10 Jahren die Entwicklung von Software leiten?
		}
	\end{enumerate}
}

\section{Course Overview}
\input{content/01c-course}
\inputuniversity{content/01c-course}
\lessonslearned{
	\item How is this course organized?
	\item Next Lecture: Why do programming languages matter?
}{
	\item Main Book: \sommervillelink{Ian Sommerville. Software Engineering, 10. Edition, Pearson, 2018.}
}{
	\begin{itemize}
		\item Any questions?
	\end{itemize}
}

%\faq{
%	\item
%}{
%	\item
%}{
%	\color{gray}\item What you should know?
%	\item What you will learn?
%	\item What you might need?
%%	\item Who are the authors of this course?
%	\item How is this course organized?
%}

\input{template/footer}
