\ifuniversity{tubs}{\date{April 14, 2026}}

\author{Thomas Thüm}
\lecture{Evolution}{evolution}

\section{Laws on Software Evolution}
\input{content/02a-laws}
\lessonslearned{
	\item Incremental development and changed requirements cause evolution
	\item Lehman's laws: software is modified and gets more complex\\
	Continuing Change, Increasing Complexity, Conservation of Familiarity, Continuing Growth, Declining Quality, Feedback System
	\item Next: How to reduce the complexity?
}{
	\item \sommerville\mychapter{2.3 Coping with Change}
	\item \sommerville\mychapter{4.6 Requirements Change}
	\item \sommerville\mychapter{9 Software Evolution}
	\item \lehmanslaws
	%\item \ludewiglichter, 23.1 \deutsch{Software-Evolution}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Try to find a (counter) example for each of Lehman's laws
		\item Discuss the example with your colleagues -- does it fulfill further laws?
	\end{enumerate}
}

\section{Smells and Refactorings}
\input{content/02b-debt}
\lessonslearned{
	\item Smells / antipatterns: structures to avoid
	\item Examples: mysterious name, duplicated code, long method
	\item Refactorings: clean-up of structures before and after every change
	\item Examples: rename X refactoring, extract/pull-up method refactoring
	\item Next: What are principles against smells?
}{
	\item \refactoring
	\item \cleancode
	% book on antipatterns
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item What further smells and refactorings do you know?
		\item Enter these smells and refactorings using the link in \StudIP

		      ~

		      \centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/plugins.php/vipsplugin/sheets/show_assignment?cid=635f5186b364a369979d0f45ab2ace1d&assignment_id=63485}
	\end{enumerate}
}

\section{Principles for Software Evolution}
\input{content/02c-principles}
\lessonslearned{
	\item KISS --- Keep It Simple, Stupid!
	\item DRY --- Don't Repeat Yourself
	\item RERO --- Release Early, Release Often
	\item Next: How to maintain software?
}{
	\item \thepragmaticprogrammer
}{
	Quiz in \StudIP

	~

	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/plugins.php/vipsplugin/sheets/show_assignment?cid=635f5186b364a369979d0f45ab2ace1d&assignment_id=63474}
}

\faq{
	\item What are reasons to change existing software?
	\item What are alternative to software changes?
	\item What activities are involved when changing software?
	\item What is the law of \ldots continuing change, increasing complexity, conservation of familiarity, continuing growth, decling quality, feedback system?
	\item What follows from Lehman's laws?
	\item What are (counter) examples for Lehman's laws?
}{
	\item Why does the readibility of code matter?
	\item What is Eagleson's law?
	\item Why are go-to's problematic?
	\item What is spaghetti or lasagna code?
	\item What is the boy scouts rule and what are its implications for software engineering?
	\item How does a child-rearing philosophy help with software development?
	\item What is the motivation for refactorings?
	\item What are (examples for) refactorings and code smells?
	\item What is a mysterious name, code clone, duplicated code, or long method?
	\item What is a rename, extract method, or pull-up method refactoring?
}{
	\item Why can it be productive to remove code?
	\item What is the KISS, DRY, RERO principle?
	\item Which refactorings or smells are related to those principles and why?
	\item Which principle is violated in a given example code?
	\item What is better? Correct or clean or consise or fast code?
}

\input{../se1/template/footer}

% TODO L03 Maintenance

\ifuniversity{tubs}{\date{April 20, 2026}}

\author{Thomas Thüm}
\lecture{Maintenance}{maintenance}

\section{Kinds of Maintenance}
\input{content/03a-maintenance}
\lessonslearned{
	\item Maintenance (Phase)
	\item Kinds: adaptive, corrective, perfective, preventive
	\item Examples for maintenance
	\item Maintenance vs evolution
	\item Next: What are common reengineering task?
}{
	\item \ludewiglichter\mychapter{22 \deutsch{Software-Wartung}}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students for Think-Pair-Share
		\item Think (2 min): find examples for two maintenance categories
		\item Pair (5 min): explain example to your group and let colleagues guess its category
		\item Share: what examples were hard to classify?
	\end{enumerate}
}

\section{Reengineering}
\input{content/03b-reengineering}
\lessonslearned{
	\item Reengineering =
	\item reverse engineering +
	\item refactoring +
	\item forward engineering
	\item Next: How to migrate legacy software?
}{
	\item \ludewiglichter\mychapter{23 \deutsch{Reengineering}}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students for Think-Pair
		\item Think (2 min): find examples for two reengineering categories
		\item Pair (5 min): explain example to your group and let colleagues guess its category
	\end{enumerate}
}

\section{Migration of Legacy Systems}
\input{content/03c-migration}
\lessonslearned{
	\item Legacy software: typical properties, examples
	\item Software migration
	\item Migration strategies: big bang migration, incremental migration, wrapping, redevelopment
	\item Next: How to design systems simplifying particular future changes?
}{
	\item \ludewiglichter\mychapter{23 \deutsch{Reengineering}}
}{
	Quiz in \StudIP

	~

	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421046}
}

\faq{
	\item What is the operations and maintenance phase?
	\item What is software maintenance and how is it different from evolution?
	\item What are the different kinds of maintenance?
	\item What are (examples for) corrective, perfective, adaptive, preventive maintenance?
	\item Given an example maintenance activity, what kind(s) of maintenance is it?
	\item Give examples that belong to several kinds of maintenance!
}{
	\item What is forward engineering, reverse engineering, restructuring, reengineering?
	\item Give examples for the above mentioned reengineering task!
	\item Given an example, classify what kind of reengineering it is!
	\item Explain the difference between X and Y by means of examples!
}{
	\item What is legacy software?
	\item What are typical properties of legacy software?
	\item Given an example description, argue whether it is legacy software or not!
	\item What is software migration?
	\item What are migration strategies?
	\item Given an example migration, identify which migration strategy was applied!
	\item Give an example where migration X is better suited than Y!
	\item What is the end of service/life?
	\item What are reasons to discontinue software projects?
}

\input{../se1/template/footer}


\ifuniversity{tubs}{\date{May 19, 2025}}

\author{Florian Sihler}
\lecture{Static Analysis}{staticanalysis}

\begin{refsection}[references/06-sa-references.bib]
\begingroup
\setcounter{tocdepth}{2}
\section{A Primer in Static Analysis}
%\input{content/a-}
\input{content/06a-sa-intro}
\lessonslearned{
	\item What is the basic idea of static analysis?
	\item How do the approaches differ?
	\item Next: How does static analysis work?
}{
	\item \fullcite{rival2020introduction}
	\item \fullcite{DBLP:journals/ftpl/Mine17}
}{
	So far:
	\begin{enumerate}
		\item<+-> Feedback, comments, or questions?
		\item<+-> Which perspective do you think is used for what? 
	\end{enumerate}
}


% \lessonslearned{
% 	\item ?
% 	\item Next: ?
% }{
% 	\item \sommerville\mychapter{?} 
% }{
% 	\begin{enumerate}
% 		\item<+-> Form groups of 2--3 students
% 		\item<+-> ?
% 	\end{enumerate}
% }

\section{Theoretical Foundations}
\input{content/06b-sa-theory}
\lessonslearned{
	\item What is a property, a lattice, a domain?
	\item Why are we interested in fixpoints?
	\item What is Rice's theorem?
	\item What are soundness and completeness?
	\item Next: What are static analyses in the real world?
}{
	\item \fullcite{cousout2021principles}
}{
	\begin{enumerate}
		\item<+-> Let's derive the Interval Lattice! \begin{itemize}
			\item What are \(\top\) and \(\bot\)?
			\item How can we define the partial order \(\sqsubseteq\)?
			\item How to define \(\sqcup\) and \(\sqcap\)?
		\end{itemize}
	\end{enumerate}
}

\section{Static Analysis in the Real-World}
\input{content/06c-sa-real-world}
\lessonslearned{
	\item What is static analysis used for in the real world?
	\item How do theory and practice differ?
	\item Next: What are dynamic analyses?
}{
	\item \fullcite{10.5555/1592955}
	\item \fullcite{DBLP:conf/pldi/FerraraNAC21}
}{
	\begin{enumerate}
		\item<+-> What tools do you use?
		\item<+-> Have you experienced problems with false positives?
		\item<+-> How to handle false positives? In what scenarios are they harmful?
	\end{enumerate}
}

\begin{frame}[allowframebreaks]{References --- Static Analysis}
\renewcommand*{\bibfont}{\tiny}%
\urlstyle{same}
\printbibliography[heading=none]
\end{frame}
\endgroup
\end{refsection}

\faq{
	\item What is static analysis and what does it do?
	%\item What is the central mathematical background of static analysis?
	\item What is Rice's theorem and why is it critical for static analysis?
	\item What are common types of static analyses?
	\item What are common perspectives of software programs used in static analysis? 
	%\item What are commonly used analyses used for static analysis?
}{
	\item What is abstract interpretation?
	\item What are (examples for) dataflow, safety, and numerical properties of programs?
	\item What is trace abstraction and why is it necessary?
	%\item What is a property and how can we describe it?
	%\item What are partial orders on sets and why are they important for static analyses?
	\item What is (an example for) an abstract domain or a lattice?
	\item Given an abstract domain (or lattice), what is the corresponding lattice (or abstract domain)?
	%\item What is a chain and why are greatest lower bounds and least upper bounds important?
	%\item What are abstract domains and what do they have to do with lattices?
	\item What is a sign analysis and how does it work?
	%\item What are common abstract domains and what are they used for?
	\item Given a couple of assignments, what is their value in an abstract domain?
	%\item How does one use a abstract domain for a simple program analysis?
	%\item What are fixpoints and why are they helpful for abstract interpretation?
	\item What is the confusion matrix of static analyses?
	\item What are precision, recall, soundness, completeness? 
}{
	%\item What are common capabilities offered by static analysis tools?
	%\item How do static analysis tools operate?
	\item What are examples of common static analysis tools?
	%\item How do exemplary static analysis tools (e.g., SonarLint, flowR) operate?
	%\item What advantages does the LiSA framework offer and how does one use it?
	\item What is an analysis for unused assignments, hardcoded credentials, intervals?
	\item How does it work?
}
