% TODO L10 Open-Source Software

\ifuniversity{tubs}{\date{June 23, 2025}}

\author{Thomas Thüm}
\lecture{Open-Source Software}{opensource}

% TODO add example: https://www.computing.co.uk/news/2025/open-source/denmark-to-replace-microsoft-office-with-open-source-alternatives

\section{Open-Source Software}
\input{content/10a-opensource}
\lessonslearned{
	\item Open-source development and open-source software
	\item Examples and central questions around open source
	\item Free software foundation: four freedoms, RMS, FOSS
	\item Next: What are main principles of open-source software?
}{
	\item \sommerville\mychapter{7.4} Open-Source Development
	\item \href{https://www.youtube.com/watch?v=Ag1AKIl_2GM}{Stallman's 2014 TEDx Talk}
%\href{https://www.ulm.ccc.de/ccc/chaosseminar/2004_11_gefahren_der_softwarepatente/}{Stallman's lecture in Ulm}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Which open-source software are you using?
		\item Have you ever made use of the four freedoms?
	\end{enumerate}
}

\section{Open-Source Principles}
\input{content/10b-principles}
\lessonslearned{
	\item Need for and impact of software licenses
	\item Ownership and code without any license
	\item Recommendations for companies
	\item Copyleft and its impact on software
	\item Next: What are major licenses?
}{
	\item \sommerville\mychapter{7.4} Open-Source Development
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Does published code without a license fulfill the four freedoms?
		\item What is the problem of creating a license for every project?
	\end{enumerate}
}

\section{Open-Source Licenses}
\input{content/10c-licenses}
\lessonslearned{
	\item Software licenses: MIT, GPL, LGPL
	\item License incompatibilities: desired and undesired
	\item Re-licensing and dual licensing
	\item Next: How to systematically reuse software?
}{
	\item \sommerville\mychapter{7.4} Open-Source Development
}{
	Quiz in \StudIP

	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421053}
}
% se1 interaction
%	\begin{enumerate}
%		\item Form groups of 2--3 students
%		\item Find three example for software systems were you would recommend any of the following licenses: MIT, GPL, LGPL
%		\item What are reasons for the license over the other two licenses?
%	\end{enumerate}

\faq{
	\item What characterizes open-source development?
	\item What are examples for successful open-source projects?
	\item What are the central questions on the usage of open-source software in a software project?
	\item What are Stallman's four freedoms?
	\item What is free software and how does it differ to freeware?
	\item What is the Free Software Foundation?
}{
	\item With whom lies code ownership and why does this necessitate software licenses?
	\item What should companies keep in mind when using open-source software?
	\item What can companies do to keep used open-source projects alive?
	\item What is the difference between copyleft and copyright? 
	\item What can mistakes on software licenses cause?
}{
	\item What are examples for common open-source licenses?
	\item Which restrictions does the MIT license have?
	\item What are the GPL and LGPL licenses and how are they related?
	\item Why are some licenses incompatible and others not?
	\item Which license is usable in which usage scenario?
	\item What is re-licensing and which requirements does it have?
	\item What is dual-licensing (used for)?
}

\input{../se1/template/footer}

% TODO L11 Software Product Lines

\ifuniversity{tubs}{\date{June 30, 2025}}

\author{Thomas Thüm}
\lecture{Software Product Lines}{productlines}

\section{Modeling of Product Lines}
\input{content/11a-modeling-spl}
\lessonslearned{
	\item Motivation for mass customization
	\item Software product lines, features, dependencies between features, configurators
	\item Feature models: abstract and concrete features, tree and cross-tree constraints
	\item Tree constraints: optional, mandatory, or group, alternative group
	\item Next: How to implement the variability specified in a feature model?
}{
	\item \fospl, Chapter~1 Software Product Lines and Section~2.3 Feature Modeling
}{
	\begin{enumerate}
		\item Model the domain of USB cables and adapters as a feature model (cf. examples shown in the lecture)
		\item Ask a fellow student to check whether your model is correct (syntax of feature model + examples are valid configurations)
		\item Make corrections to your feature model if needed
		\item (Keep feature model for later task)
	\end{enumerate}
}

\section{Implementation of Product Lines}
\input{content/11b-implementing-spl}
\lessonslearned{
	\item Branches, runtime parameters
	\item Preprocessors: in-place or not, feature traceability
	\item Frameworks with plug-ins
	\item Next: How to test the modeled and implemented variability?
}{
	\item \featureide\ Chapter~9 Conditional Compilation with FeatureIDE and Chapter~17 Tool Support Beyond Preprocessors and Feature Modules
}{
	\begin{enumerate}
		\item Give your feature model to a fellow student
		\item Check the feature model of a fellow student:\\how many example products are modeled?
	\end{enumerate}
}

\section{Testing of Product Lines}
\input{content/11c-testing-spl}
\lessonslearned{
	\item All configurations? one configuration? feature interactions!
	\item Combinatorial interaction testing:\\pairwise and t-wise
	\item Trade-off between:\\sample coverage (effectiveness) vs\\ sample size (testing efficiency) vs\\ time to compute sample (sampling efficiency)
	\item Next: Why is automotive software so challenging?
}{
	\item \icpl --- famous ICPL algorithm
	\item \href{https://dl.acm.org/doi/10.1145/3712193}{Krupke~et~al.} --- first algorithm that can compute the smallest sample
}{
	Quiz in \StudIP

	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421054}
}

\faq{
	\item What is mass customization, a software product line, a feature?
	%\item Which advantages do software product lines offer and which problems do they solve?
	\item Why are there dependencies between features?
	\item What is feature modeling?
	\item How are feature models defined?
	\item How can we use feature models to model (software) variability?
	\item What is the problem when enumerating all valid configurations of a software product line?
}{
	\item What are the advantages of feature modeling in software development? 
	\item How can we configure and implement runtime variability, runtime parameters, runtime properties?
	\item How can we implement static variability?
	\item What kind of preprocessors are there and how do they differ?
	\item What are (dis)advantages of preprocessors?
	\item How to implement variability with frameworks/plug-ins?
}{
	\item How can we test variable software?
	\item What are (dis)advantages of testing all configurations or only one configuration?
	\item What are feature interactions and why are they challenging?
	\item What is pairwise and t-wise interaction testing?
	\item What are (dis)advantages of pairwise interaction testing?
	\item What is the fundamental trade-off in t-wise interaction testing?
	\item How does t-wise interaction testing scale, in both sample size and computation time?
	\item How to reduce the potential for feature interactions?
}

\input{../se1/template/footer}

% TODO L12 Automotive Software

\ifuniversity{tubs}{\date{July 7, 2025}}

\author{Thomas Thüm}
\lecture{Automotive Software}{automotive}

\begin{frame}[plain]{Book Used for this Lecture \mytitlesource{\staron}}
	\centering\staronlink{\pic[height=\textheightwithtitle,page=1]{automotive/S21}}
\end{frame}

\section{Automotive Systems}
\input{content/12a-automotive}
\lessonslearned{
	\item Software in automotive systems
	\item with insufficient quality
	\item with its characteristics and complexity
	\item Next: How can software architecture help?
}{
	\item \staronlink{Miroslaw Staron, Automotive Software Architectures, Chapter~1 (p.~vii--18)}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item How have software failures in cars affected you?
		\item How were the issues solved?
	\end{enumerate}
}

\section{Automotive Software Architectures}
\input{content/12b-automotivearchitectures}
\lessonslearned{
	\item Federated and centralized software architectures
	\item Standardization of architectures with AUTOSAR
	\item Example application of software architectures
	\item Next: What were the topics of this course?
}{
	\item \staronlink{Miroslaw Staron, Automotive Software Architectures, Chapter~2 (p.~19--53)}
	\item \staronlink{Miroslaw Staron, Automotive Software Architectures, Chapter~3 (p.~55--66)}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Are federated or centralized architectures better suited \ldots{}
		\begin{itemize}
			\item to implement autonomous driving?
			\item for over-the-air updates?
		\end{itemize}
	\end{enumerate}
}

\section{Course Summary}
\input{content/12c-summary}
\lessonslearned{
	\item How to \emph{change} software?
	\item How to \emph{verify} software?
	\item How to \emph{reuse} software?
	\item Next: What are experiences of applying SE principles in practice?
}{
	\item see other lectures
}{
	Questions?
}

%\faq{
%	\item
%}{
%	\item
%}{
%	\item
%}

\input{../se1/template/footer}
