% !TeX spellcheck = en_US

% TODO L05 Compilation

\ifuniversity{tubs}{\date{May 12, 2025}}

\author{Thomas Thüm}
\lecture{Compilation}{compilation}

\section{Fundamentals on Compilation}
\input{content/05a-compilation}
\lessonslearned{
	\item 2020 Turing Award for high-level programming languages
	\item Compilation vs interpretation
	\item Type safety and correctness
	\item Next: How do compilers work?
}{
	\item \dragonbook
}{
	\begin{enumerate}
		\item Form groups of 2--3 students for Think-Pair-Share
		\item What are chances and risks of high-level programming languages (i.e., an increasing gap between high-level instructions and low-level machine code)?
	\end{enumerate}
}

\section{Architecture of Compilers}
\input{content/05b-compilerarchitecture}
\lessonslearned{
	\item Compiler architecture
	\item Compile time vs run time: goals of optimizations
	\item Just-in-time compilation
	\item Compiler optimizations
	\item Intermediate languages
	\item Next: What are common misunderstandings of developers related to compilation?
}{
	\item \dragonbook
}{
	Quiz in \StudIP

	~

	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421048}
}
\begin{frame}[fragile]{Illustration of Compiler Optimizations}
	\begin{fancycolumns}
		\begin{example}{Example Java Program}
			\begin{lstlisting}[style=java]
class A {
	public static void main(String[] args) {
		A x = new B();
		Object y = new String();
		System.out.print(x.m(y));
	}
	int m(Object o) { return 1; }
	int m(String s) { return 2; }
}
class B extends A {
	int m(Object o) { return 3; }
	int m(String s) { return 4; }
}
			\end{lstlisting}
		\end{example}
		\nextcolumn
		\uncover<2->{\begin{example}{What is the output?}
			\begin{enumerate}
				\item program as shown left % 3
				\item if we change type of x to B % 3
				\item if we change type of y to String % 4
				\item if we remove method m(Object o) in class B % 1
				\item if we remove method m(Object o) in class A % undefined
			\end{enumerate}
		\end{example}}
		\uncover<3->{\begin{note}{Hint}
			Java uses dynamic dispatch for overriding (e.g., x)\\ and static dispatch for overloading (e.g., y)
			
			~
			
			the former is required for inheritance, the latter a design decision of Java
		\end{note}}
	\end{fancycolumns}
\end{frame}

\section{Misunderstandings on Performance}
\input{content/05c-misunderstandings}
\lessonslearned{
	\item What are common misunderstandings about performance?
	\item Why are array, loops, method calls, objects, and garbage collection slow?
	\item What is the connection between compiler optimizations, smells, and refactorings?
	\item Next: How do static analyses work?
}{
	\item \dragonbook
}{
	Content of last part rather subjective as based on Thomas' experiences.
	
	~
	
	Any feedback or comments or questions?
}

\faq{
	\item What are examples for modern programming languages?
	\item Why are there multiple programming languages today?
	\item What is the difference between source code compilation and interpretation?
	\item What are errors and warnings in the context of compilation?
	\item How do we categorize errors?
	\item What are lexical errors and how do we fix them?
	\item What are syntax errors and how do we fix them? 
	\item What are type errors and how do we fix them?
	\item What are runtime errors and how do we fix them?
	\item What are logical errors and how do we fix them?  
	\item What is the difference between type safety and type correctness?
}{
	\item What is a pipe-and-filter architecture and what problems does it address?
	\item What is the Chomsky hierarchy?
	\item What impact does the Chomsky hierarchy have on compilers?
	\item What are the common stages in a compiler?
	\item What are compiler optimizations and what do they do?
	\item What is the difference between run time and compile time
	\item What are advantages and disadvantages of just-in-time compilation?
	\item What are common compiler optimizations and in which stages do they operate?
	\item What are advantages of intermediate languages? 
}{
	\item What are common misunderstandings in the context of program optimization?
	\item Why can array access be slow and why is this not important?
	\item Why can loops be slow and why is this not important?
	\item Why do method calls generate overhead and why is this not important?
	\item Why are objects slow and why is this not important?
	\item Why is garbage collection slow and why is this not important?
	\item What is the connection between compiler optimizations, smells, and refactorings?
	\item Why are not all instructions equally fast?
	\item What is the golden rule of program optimization?
	\item Why is simplicity typically more important than performance?
}

\input{../se1/template/footer}

% TODO L06 Static Analysis

\ifuniversity{tubs}{\date{May 19, 2025}}

\author{Florian Sihler}
\lecture{Static Analysis}{staticanalysis}

\begin{refsection}[references/06-sa-references.bib]
\begingroup
\setcounter{tocdepth}{2}
\section{A Primer in Static Analysis}
%\input{content/a-}
\input{content/06a-sa-intro}
\lessonslearned{
	\item What is the basic idea of static analysis?
	\item How do the approaches differ?
	\item Next: How does static analysis work?
}{
	\item \fullcite{rival2020introduction}
	\item \fullcite{DBLP:journals/ftpl/Mine17}
}{
	So far:
	\begin{enumerate}
		\item<+-> Feedback, comments, or questions?
		\item<+-> Which perspective do you think is used for what? 
	\end{enumerate}
}


% \lessonslearned{
% 	\item ?
% 	\item Next: ?
% }{
% 	\item \sommerville\mychapter{?} 
% }{
% 	\begin{enumerate}
% 		\item<+-> Form groups of 2--3 students
% 		\item<+-> ?
% 	\end{enumerate}
% }

\section{Theoretical Foundations}
\input{content/06b-sa-theory}
\lessonslearned{
	\item What is a Property, a Lattice, a Domains?
	\item Why are we interested in Fixpoints?
	\item What is Rice's Theorem?
	\item What are Soundness and Completeness?
	\item Next: Static Analysis in the Real World
}{
	\item \fullcite{cousout2021principles}
}{
	\begin{enumerate}
		\item<+-> Let's derive the Interval Lattice! \begin{itemize}
			\item What are \(\top\) and \(\bot\)?
			\item How can we define the partial order \(\sqsubseteq\)?
			\item How to define \(\sqcup\) and \(\sqcap\)?
		\end{itemize}
	\end{enumerate}
}

\section{Static Analysis in the Real-World}
\input{content/06c-sa-real-world}
\lessonslearned{
	\item What is Static Analysis used for in the Real World?
	\item How do Theory and Practice differ?
	\item Next: Dynamic Analysis
}{
	\item \fullcite{10.5555/1592955}
	\item \fullcite{DBLP:conf/pldi/FerraraNAC21}
}{
	\begin{enumerate}
		\item<+-> What tools do you use?
		\item<+-> Have you experienced problems with false positives?
		\item<+-> How to handle false positives? In what scenarios are they harmful?
	\end{enumerate}
}

\begin{frame}[allowframebreaks]{References --- Static Analysis}
\renewcommand*{\bibfont}{\tiny}%
\urlstyle{same}
\printbibliography[heading=none]
\end{frame}
\endgroup
\end{refsection}

\faq{
	\item What is static analysis and what does it do?
	\item What is the central mathematical background of static analysis?
	\item What are common types of static analyses and how do they differ?
	\item What are common perspectives of software programs used in static analysis? 
	\item What are commonly used analyses used for static analysis?
}{
	\item What is abstract interpretation?
	\item Which properties of software programs are typically analyzed?
	\item What is trace abstraction and why is it necessary?
	\item What is a property and how can we describe it?
	\item What are partial orders on sets and why are they important for static analyses?
	\item What is a complete lattice and what are common examples?
	\item What is a chain and why are greatest lower bounds and least upper bounds important?
	\item What are abstract domains and what do they have to do with lattices?
	\item What are common abstract domains and what are they used for?
	\item How does one use a abstract domain for a simple program analysis?
	\item What are fixpoints and why are they helpful for abstract interpretation?
	\item What is Rice's theorem?
	\item What is a confusion matrix, precision and recall?
	\item How are soundness and completeness fulfilled in the context of abstract interpretation? 
}{
	\item What are common capabilities offered by static analysis tools?
	\item How do static analysis tools operate?
	\item What are examples of common static analysis tools?
	\item How do exemplary static analysis tools (e.g., SonarLint, flowR) operate?
	\item What advantages does the LiSA framework offer and how does one use it?
}


\input{../se1/template/footer}

% TODO L07 Dynamic Analysis

\ifuniversity{tubs}{\date{May 26, 2025}}

\author{Raphael Dunkel}
\lecture{Dynamic Analysis}{dynamicanalysis}

\section{Dynamic Analysis -- Recap}
\input{content/07a-testingrecap}
\lessonslearned{
	\item Why is software quality assurance important?
	\item What can we do to improve software quality?
	\item What types of tests do exist?
	\item Next: How do we test parts of our application?
}{
	\item \sommerville\mychapter{8} 
	\item \ludewiglichter, Chapter 5 and Chapter 13
}{
	\begin{enumerate}
		\item<+-> Form groups of 2--3 students
		\item<+-> Try to answer all questions on the left and discuss them with your partners
	\end{enumerate}
}

\section{Unit and Integration Testing}
\input{content/07b-unitintegration}
\lessonslearned{
	\item How and why do we perform unit tests?
	\item What are common tools to assist with unit testing?
	\item Why do we perform integration tests?
	\item What is coverage and why is it important?
	\item Next: How do we test the whole application?
}{
	\item \sommerville\mychapter{8} 
	\item \cohnagile
}{
	\begin{enumerate}
		\item<+-> Form groups of 2--3 students
		\item<+-> Think about the CrowdStrike and Therac-25 incidents. Discuss whether unit and integration tests could have prevented both incidents
		\item<+-> How would you have tested both systems?
	\end{enumerate}
}

\section{System and Acceptance Testing}
\input{content/07c-systemacceptance}
\lessonslearned{
	\item What is end-to-end testing?
	\item What is fuzzing?
	\item Why do we still use manual tests?
	\item What are the steps to program acceptance?
	\item Next: Are there other ways to prevent software errors?
}{
	\item \sommerville\mychapter{8} 
}{
	\begin{enumerate}
		\item<+-> Form groups of 2--3 students
		\item<+-> Think about the CrowdStrike and Therac-25 incidents. Discuss whether system and acceptance tests could have prevented both incidents
		\item<+-> How would you have tested both systems?
	\end{enumerate}
}

\faq{
	\item Why do we need software quality assurance?
	\item What can follow from missing software quality assurance?
	\item What is software quality?
	\item What types of testing are there?
	\item How do we assure software quality?
	\item What are systematic tests and test cases?
	\item Which types of testing exist and why are there multiple?
	\item What are the stages of testing and how do they differ?
}{
	\item Why and how do we automate software testing?
	\item What are test doubles and why do we use them?
	\item How is the quality of a test suit measured?
	\item What are flaky tests and how do we handle them?
	\item How do we test component interactions?
	\item Why do we need different kinds of code coverage?
}{ 
	\item What are end-to-end tests?
	\item Why do we still employ manual tests?
	\item How are systematic manual tests structured?
	\item Why and how do we use randomized program input for security testing?
	\item What is acceptance testing?
	\item What kinds of user tests are there?
	\item What happens before a software product is accepted by the customer?
	\item How do we test graphical user interfaces?
}

\input{../se1/template/footer}

% TODO L08 Design by Contract

\ifuniversity{tubs}{\date{June 2, 2025}}

\author{Thomas Thüm}
\lecture{Design by Contract}{designbycontract}

\section{Documentation of Assumptions}
\input{content/08a-assumptions}
\lessonslearned{
	\item Why did Ariane 5 explode?
	\item What could have prevented the failure?
	\item How can developers document their assumptions?
	\item Code documentation, defensive programming, runtime assertions, unit testing
	\item Why is none of these techniques superior?
	\item Next: How to write unabiguous and accessible specifications?
}{
	\item \lessonsofariane % TODO more literature pointers would be helpful
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Student explains a column to their group, repeat with other students/columns
		\item Try to answer the three questions (resulting in new columns)
	\end{enumerate}
}

\section{Design by Contract with JML}
\input{content/08b-designbycontract}
\lessonslearned{
	\item Design by contract
	\item Behavioral interface specification languages
	\item Java Modeling Language (JML): requires, ensures, invariant, assignable, \ldots
	\item Next: How does design by contract work beyond JML?
}{
	\item \hatcliff{} --- survey on behavioral interface specification languages
	\item \purity{} --- different notions of pure methods / methods without side-effects % TODO could also be a good extension to the slides to add Figure 1 and Table 3
}{
	\begin{itemize}
		\item Form groups of 2--3 students
		\item Discuss what you learned so far and try to apply that knowledge
		\item What are advantages and disadvantages of ...?
			\begin{enumerate}
				\item class invariants
				\item assignable clauses
				\item method calls in contracts
			\end{enumerate}
	\end{itemize}
}

\section{Design by Contract with SPARK}
\input{content/08c-spark}
\lessonslearned{
	\item Ada vs SPARK
	\item Design by contract with SPARK
	\item Live demo with examples
	\item Next: How is quality assurance embedded in continuous integration?
}{
	\item \introtoada
	\item \introtospark
}{
	Modify, run, and formally verify example programs with SPARK contracts

	~

	\centering\fancyqr{color=black,height=30mm}{https://learn.adacore.com/courses/intro-to-spark/chapters/03_Proof_Of_Program_Integrity.html\#contracts}
}

%\faq{
%	\item
%}{
%	\item
%}{
%	\item
%}

\input{../se1/template/footer}

% TODO L09 Configuration Management

\ifuniversity{tubs}{\date{June 16, 2025}}

\author{Raphael Dunkel}
\lecture{Configuration Management}{configuration}

\section{Configuration Management}
\input{content/09a-configmanagement}
\lessonslearned{
	\item What is Configuration Management?
	\item Why do we use VCS, automated build systems, and change management software?
	\item Why do we release differently in CI/CD vs. classical approaches?
	\item Next: How do we use version control systems?
}{
	\item \sommerville\mychapter{25} 
}{
	\begin{enumerate}
		\item<+-> Form groups of 2--3 students
		\item<+-> Pick a software system you yourself use (easiest: open source)
		\item<+-> Research which VCS, system building tool and change management process  is used
	\end{enumerate}
}

\section{Version Control Systems}
\input{content/09b-vcs}
\lessonslearned{
	\item What are version control systems?
	\item What is git?
	\item How do I use it?
	\item Next: What can we automate with VCSs?
}{
	\item \sommerville\mychapter{25} 
	\item Chacon, Straub: \href{https://git-scm.com/book/en/v2}{Pro Git}
}{
	Learn Git Branching
	
	~
	
	\centering\fancyqr{color=black,height=30mm}{https://learngitbranching.js.org/?locale=de\_DE}
}

\section{CI/CD}
\input{content/09c-cicd}
\lessonslearned{
	\item What are DevOps?
	\item What is CI/CD?
	\item Why do we do it?
	\item Next: How to reuse software (that is already tested)?
}{
	\item \sommerville\mychapter{25} 
}{
		Quiz in \StudIP
	
	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421052}
}

\faq{
	\item What is configuration management?
	\item What are the four central activities in configuration management?
	\item What are the development stages of a software product?
	\item What is a version control system?
	\item What are the different approaches to version control and how do they differ?
	\item What is system building?
	\item What functionalities does a build system offer?
	\item What is change management?
	\item How can the process of software changes be structured?
	\item What is release management?
	\item What are important steps of a software release process?
	\item What is typically contained in a software release?
}{
	\item What is distributed version control?
	\item What is git and how does it work?
	\item What are the most common git commands and what do they do?
	\item What are branches and merges?
	\item How can one solve merge conflicts?
	\item What are other ways of combining diverging development branches?
	\item What are common approaches for git branching?
}{
	\item What is DevOps?
	\item How is the DevOps cycle structured?
	\item What is system building?
	\item What is continuous integration and what are common steps?
	\item How are CI/CD pipelines structured?
	\item What is containerization?
	\item What are advantages and disadvantages of software containers?
	\item What is test-driven developments?
	\item What are common problems encountered in test-driven development and how can they be solved?
}

\input{../se1/template/footer}
