% TODO L01 Introduction

\ifuniversity{tubs}{\date{April 8, 2025}}

\author{Thomas Th端m}
\lecture{Introduction}{introduction}

\section{Recap on Software Engineering 1}
\input{content/01a-recap}
\lessonslearned{
	\item How to manage / structure software development?
	\item What software is required?
	\item How to model / design / implement / test / change / maintain software?
	\item Next: What are motivating examples for SE2?
}{
	\item[] See literature pointers in SE1
}{
	Quiz in \StudIP:
	
	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421044}
}

\section{Motivating Examples}
\input{content/01b-examples}
\lessonslearned{
	\item software failures are everywhere:\\Saxony election, CrowdStrike outage, Ariane 5 failure, Xiaomi SU7 crash
	\item and may have severe consequences
	\item software engineering aims to reduce them
	\item good software engineering practices required by law: Switzerland's government
	\item Next: How is SE2 organized?
}{
	\item[] see links on previous slides
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Familiarize with each other: name, field of study, term, other common courses, \ldots
		\item Do you know any further examples of software glitches?
		\item Could better software engineering have prevented those?
	\end{enumerate}
}

\section{Course Organization}
\input{content/01c-course}
\inputuniversity{content/01c-course}
\lessonslearned{
	\item What you: should know, will learn, might need
	\item The course: lectures, exercises
	% TODO \item Next Lecture: 
}{
	\item[] Main Book: \sommervillelink{Ian Sommerville. Software Engineering, 10. Edition, Pearson, 2018.}
}{
	\begin{itemize}
		\item Any questions?
		\item Optional: repeat SE1 quizzes from last term, all collected in one quiz for SE2 in \StudIP

			~

			\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421044}
	\end{itemize}
}

\faq{
	\item Recap on SE1
	\item How is software engineering different from programming?
	\item How are software projects managed with Gantt charts and network diagrams?
	\item How are software projects structured with process models (e.g., Waterfall model, V model)?
	\item What are functional and non-functional requirements?
	\item How to model software with UML (incl. high-level and low-level structure)?
	\item How to implement and test software?
	\item How to change software and control changes?
}{
	\item How can a software glitch effect elections?
	\item How can politians enforce the use and development of open-source software?
	\item How can a software update cause the largest IT outage?
	\item How can software reuse postpone research by one year?
	\item How can advanced driver assistance crash a car?
	% TODO Raphael: collect other examples throughout the course and add all here
}{
	\item What you should know before taking this course?
	\item What you will learn during this course?
	\item Which book is recommended for this course?
	\item How is this course organized?
}

\input{../se1/template/footer}

% TODO L02 Evolution

\ifuniversity{tubs}{\date{April 14, 2025}}

\author{Thomas Th端m}
\lecture{Evolution}{evolution}

\section{Laws on Software Evolution}
\input{content/02a-laws}
\lessonslearned{
	\item Incremental development and changed requirements cause evolution
	\item Lehman's laws: software is modified and gets more complex\\
		Continuing Change, Increasing Complexity, Conservation of Familiarity, Continuing Growth, Declining Quality, Feedback System
	\item Next: How to reduce the complexity?
}{
	\item \sommerville\mychapter{2.3 Coping with Change}
	\item \sommerville\mychapter{4.6 Requirements Change}
	\item \sommerville\mychapter{9 Software Evolution}
	\item \lehmanslaws
	%\item \ludewiglichter, 23.1 \deutsch{Software-Evolution}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item Try to find a (counter) example for each of Lehman's laws
		\item Discuss the example with your colleagues -- does it fulfill further laws?
	\end{enumerate}
}

\section{Smells and Refactorings}
\input{content/02b-debt}
\lessonslearned{
	\item Smells / antipatterns: structures to avoid
	\item Examples: mysterious name, duplicated code, long method
	\item Refactorings: clean-up of structures before and after every change
	\item Examples: rename X refactoring, extract/pull-up method refactoring
	\item Next: What are principles against smells?
}{
	\item \refactoring
	\item \cleancode
	% book on antipatterns
}{
	\begin{enumerate}
		\item Form groups of 2--3 students
		\item What further smells and refactorings do you know?
		\item Enter these smells and refactorings using the link in \StudIP

		~
		
		\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421045}
	\end{enumerate}
}

\section{Principles for Software Evolution}
\input{content/02c-principles}
\lessonslearned{
	\item KISS --- Keep It Simple, Stupid!
	\item DRY --- Don't Repeat Yourself
	\item RERO --- Release Early, Release Often
	\item Next: How to maintain software?
}{
	\item \thepragmaticprogrammer
}{
	Quiz in \StudIP

	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421045}
}

\faq{
	\item What are reasons to change existing software?
	\item What are alternative to software changes?
	\item What activities are involved when changing software?
	\item What is the law of \ldots continuing change, increasing complexity, conservation of familiarity, continuing growth, decling quality, feedback system?
	\item What follows from Lehman's laws?
	\item What are (counter) examples for Lehman's laws?
}{
	\item Why does the readibility of code matter?
	\item What is Eagleson's law?
	\item Why are go-to's problematic?
	\item What is spaghetti or lasagna code?
	\item What is the boy scouts rule and what are its implications for software engineering?
	\item How does a child-rearing philosophy help with software development?
	\item What is the motivation for refactorings?
	\item What are (examples for) refactorings and code smells?
	\item What is a mysterious name, code clone, duplicated code, or long method?
	\item What is a rename, extract method, or pull-up method refactoring?
}{
	\item Why can it be productive to remove code?
	\item What is the KISS, DRY, RERO principle?
	\item Which refactorings or smells are related to those principles and why?
	\item Which principle is violated in a given example code?
	\item What is better correct or clean or consise or fast code?
}

\input{../se1/template/footer}

% TODO L03 Maintenance

\ifuniversity{tubs}{\date{April 15, 2025}}

\author{Thomas Th端m}
\lecture{Maintenance}{maintenance}

\section{Kinds of Maintenance}
\input{content/03a-maintenance}
\lessonslearned{
	\item Maintenance (Phase)
	\item Kinds: adaptive, corrective, perfective, preventive
	\item Examples for maintenance
	\item Maintenance vs evolution
	\item Next: What are common reengineering task?
}{
	\item \ludewiglichter\mychapter{22 \deutsch{Software-Wartung}}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students for Think-Pair-Share
		\item Think (2 min): find examples for two maintenance categories
		\item Pair (5 min): explain example to your group and let colleagues guess its category
		\item Share: what examples were hard to classify?
	\end{enumerate}
}

\section{Reengineering}
\input{content/03b-reengineering}
\lessonslearned{
	\item Reengineering =
	\item reverse engineering +
	\item refactoring +
	\item forward engineering
	\item Next: How to migrate legacy software?
}{
	\item \ludewiglichter\mychapter{23 \deutsch{Reengineering}}
}{
	\begin{enumerate}
		\item Form groups of 2--3 students for Think-Pair
		\item Think (2 min): find examples for two reengineering categories
		\item Pair (5 min): explain example to your group and let colleagues guess its category 
	\end{enumerate}
}

\section{Migration of Legacy Systems}
\input{content/03c-migration}
\lessonslearned{
	\item Legacy software: typical properties, examples
	\item Software migration
	\item Migration strategies: big bang migration, incremental migration, wrapping, redevelopment
	\item Next: How to design systems simplifying particular future changes?
}{
	\item \ludewiglichter\mychapter{23 \deutsch{Reengineering}}
}{
	Quiz in \StudIP

	~
	
	\centering\fancyqr{color=black,height=30mm}{https://studip.tu-braunschweig.de/dispatch.php/course/courseware/courseware/18333?cid=635f5186b364a369979d0f45ab2ace1d\#/structural\_element/421046}
}

\faq{
	\item What is the operations and maintenance phase?
	\item What is software maintenance and how is it different from evolution?
	\item What are the different kinds of maintenance?
	\item What are (examples for) corrective, perfective, adaptive, preventive maintenance?
	\item Given an example maintenance activity, what kind(s) of maintenance is it?
	\item Give examples that belong to several kinds of maintenance!
}{
	\item What is forward engineering, reverse engineering, restructuring, reengineering?
	\item Give examples for the above mentioned reengineering task!
	\item Given an example, classify what kind of reengineering it is!
	\item Explain the difference between X and Y by means of examples!
}{
	\item What is legacy software?
	\item What are typical properties of legacy software?
	\item Given an example description, argue whether it is legacy software or not!
	\item What is software migration?
	\item What are migration strategies?
	\item Given an example migration, identify which migration strategy was applied!
	\item Give an example where migration X is better suited than Y!
	\item What is the end of service/life?
	\item What are reasons to discontinue software projects?
}

\input{../se1/template/footer}

% TODO L04 Design Patterns

\ifuniversity{tubs}{\date{April 28, 2025}}

\author{Thomas Th端m}
\lecture{Design Patterns}{designpatterns}

\newcommand{\userstoriesA}{
	\item User Story 1: Implement subtraction and division as two new operations.
	\item User Story 2: Fix that the error label does not disappear after wrong followed by correct input.
	\item User Story 3: Replace first occurrence of existing value (e.g., 3) in the formula by a new value (e.g., 5).
}
\newcommand{\userstoriesB}{
	\item User Story 4: Enable or disable debug outputs with an enum constant (no debug output, only warnings, all debug output).
	\item User Story 5: Layout all buttons in the GUI horizontally instead of vertically.
	\item User Story 6: Implement a dark mode of the calculator by means of a feature flag.
}
\newcommand{\userstoriesC}{
	\item User Story 7: Display brackets only when needed.
	\item User Story 8: If there is no error message, show seconds elapsed since last input.
	\item User Story 9: Toggle in GUI to let users choose between brackets and evaluation of subterms.
}

\section{Structural Patterns}
\input{content/04a-structuralpatterns}
\lessonslearned{
	\item Design patterns by the Gang of Four
	\item Object adapter, composite pattern, decorator pattern
	\item Next: How to solve problems with the creation of objects?
}{
	\item \gof\mychapter{4}
}{
	\begin{itemize}
		\item Download the FancyCalculator from \StudIP
		\item Install Java and IDE (Eclipse recommended)
		\userstoriesA
		\item Hint: Apply knowledge on coding conventions, smells, refactorings, and design patterns where applicable.
	\end{itemize}
}

\section{Creational Patterns}
\input{content/04b-creationalpatterns}
\lessonslearned{
	\item Singleton pattern
	\item Abstract factory pattern
	\item Next: How to solve problems with the behavior of objects?
}{
	\item \gof\mychapter{3}
}{
	Continue implementing the FancyCalculator:
	\begin{itemize}
		{\color{gray} \userstoriesA }
		\userstoriesB
	\end{itemize}
}

\section{Behavioral Patterns}
\input{content/04c-behavioralpatterns}
\lessonslearned{
	\item Observer pattern
	\item Visitor pattern
	\item Next lecture: What do software engineers need to know about compilers?
}{
	\item \sommerville\mychapter{5}
}{
	Continue implementing the FancyCalculator:\\see next slide
}

\begin{frame}
	\begin{fancycolumns}[widths={65}]
		\begin{example}{FancyCalculator - 3rd Exercise on Design Patterns}
			\begin{itemize}
				\userstoriesA
				\userstoriesB
				\userstoriesC
			\end{itemize}
		\end{example}
	\nextcolumn
		\begin{definition}{Design Patterns}
			\begin{itemize}
				\item Structural patterns:\\object adapter, composite, decorator
				\item Creational patterns:\\singleton, abstract factory
				\item Behavioral patterns:\\observer
			\end{itemize}
		\end{definition}
		\begin{note}{Hints}
			\begin{itemize}
				\item User Story 1/2: fix DRY first
				\item User Story 3: try composite
				\item User Story 4: try singleton
			\end{itemize}
		\end{note}
	\end{fancycolumns}
\end{frame}

%\faq{
%	\item
%}{
%	\item
%}{
%	\item
%}

\input{../se1/template/footer}
