% TODO L05 PROJECT MANAGEMENT

\ifuniversity{tubs}{\date{November 18, 2025}}

\author{Thomas Thüm}
\lecture{Project Management}{management}

\begin{frame}{\insertsubtitle}
	\alt<3->{%
		\projectcartoon{04}{how the programmer implemented it\\~\\\textbf{implementation}} % se02 implementation
		\projectcartoon{05}{what the beta testers received\\~\\\textbf{testing}} % se03 testing
		\projectcartoon{18}{how patches were applied\\~\\\textbf{maintenance}} % se04 maintenance
	}{%
		\hprojectcartoon{04}{how the programmer implemented it\\~\\\textbf{implementation}} % se02 implementation
		\hprojectcartoon{05}{what the beta testers received\\~\\\textbf{testing}} % se03 testing
		\hprojectcartoon{18}{how patches were applied\\~\\\textbf{maintenance}} % se04 maintenance
	}%
	\uncover<2->{\hprojectcartoon{09}{how the customer was billed\\~\\\textbf{project management}}} 
\end{frame}

\section{Introduction to Project Management}
\input{content/05a-projectmanagement}
\lessonslearned{
	\item Software development projects
	\item Project management: goals, influences, activities
	\item Risk and people management
	\item Next: How to create realistic project plans?
}{
	\item \sommerville\mychapter{22} Project Management
	\item \ludewiglichter\mychapter{7.2} \deutsch{Software-Projekte}
}{
	\begin{enumerate}
		\item<+-> Form groups of 2-3 students
		\item<+-> Risk identification and analysis: Give an example for a risk of a messenger app and specify probability, severity, and classification
		\item<+-> Risk planning and monitoring: How to address the risk mentioned by one of your colleagues? What could change during the project?
		\item<+-> Repeat 2. and 3. with other examples
	\end{enumerate}
}

\section{Project Planning and Scheduling}
\input{content/05b-planning}
\lessonslearned{
	\item Project planning and software pricing
	\item Project scheduling with Gantt charts and network diagrams
	\item Next: How to estimate the project costs?
}{
	\item \sommerville\mychapter{23} Project Planning
	\item \ludewiglichter\mychapter{8.3.2} \deutsch{Projektphasen}
}{
	Fill out the network diagram in Stud.IP
	%\lectureqr{06}
}

\section{Cost Estimation}
\input{content/05c-costestimation}
\lessonslearned{
	\item Manageable units with \emph{user stories}
	\item Effort/cost estimation with \emph{planning poker}
	\item Tracking progress with \emph{burndown charts}
	\item Cost estimation is an iterative process
	\item Next: How to manage the process beyond implementation (e.g., testing)?
}{
	\item[] \sommerville\mychapter{23} Project Planning % TODO check citation
}{
	Questions?
%	\begin{enumerate}
%		\item<+-> Form groups of 3-4 students
%		\item<+-> Choose a feature of an app of your choice
%		\item<+-> In isolation: estimate effort in weeks to develop that feature
%		\item<+-> In group: name your estimate and discuss outliers
%		\item<+-> Repeat 3--4 until consensus is reached
%	\end{enumerate}
}

%\faq{
%	\item
%}{
%	\item
%}{
%	\item
%}

\input{template/footer}

% TODO L06 DEVELOPMENT PROCESS

\ifuniversity{tubs}{\date{November 25, 2025}}

\author{Thomas Thüm}
\lecture{Development Process}{process}

\begin{frame}{The Process of Food Delivery}
	\begin{fancycolumns}
		\pic[width=\linewidth]{process/food-delivery}
		\nextcolumn
		\begin{example}{Phases of Food Delivery}
			eat, deliver, order, pay, choose, wait?
		\end{example}
	\end{fancycolumns}
\end{frame}

\begin{frame}{\insertsubtitle{} \deutsch{Vorgehensmodelle}}
%	\alt<2->{%
		\hprojectcartoon{04}{how the programmer implemented it\\~\\\textbf{implementation}} % se02 implementation
		\hprojectcartoon{05}{what the beta testers received\\~\\\textbf{testing}} % se03 testing
		\hprojectcartoon{18}{how patches were applied\\~\\\textbf{maintenance}} % se04 maintenance
		\hprojectcartoon{09}{how the customer was billed\\~\\\textbf{project management}} % se05 project management
%	}{%
%		\hprojectcartoon{18}{how patches were applied\\~\\\textbf{maintance}} % se04 maintenance
%		\hprojectcartoon{09}{how the customer was billed\\~\\\textbf{project management}} % se05 project management
%		\hprojectcartoon{05}{what the beta testers received\\~\\\textbf{testing}} % se03 testing
%		\hprojectcartoon{04}{how the programmer implemented it\\~\\\textbf{implementation}} % se02 implementation
%	}%
\end{frame}

\section{Development Processes: The~Waterfall~Model}
\input{content/06a-waterfall}
\lessonslearned{
	\item Software development process: motivation and goal
	\item Waterfall model: phases, results, example domains, advantages and disadvantages
	\item Next: How is software developed for German government?
}{
	\item[] \sommerville\mychapter{2} Software Processes
}{
	Quiz in Stud.IP
	\lectureqr{06}
}

\section{V-Model for Extensive Testing}
\input{content/06b-vmodel}
\lessonslearned{
	\item V-model
	\item Testing stages: unit testing, integration testing, system testing, acceptance testing
	\item Next: What is the most frequently used process model?
}{
	\item \ludewiglichter\mychapter{10.3} \deutsch{Das V-Modell}
	\item \sommerville\mychapter{2.2.3} Software Validation
	\item \sommerville\mychapter{8.1} Development Testing
}{
	\begin{enumerate}
		\item<+-> Form groups of 2-3 students
		\item<+-> 2 minutes of silence: Choose one of the four testing stages and think of an example of a fault that can be detected in that testing stage.
		\item<+-> 5--10 minutes of group discussions: Discuss those faults and whether they could and should have been found in the other stages.
	\end{enumerate}
}

\section{Scrum for Agile Development}
\input{content/06c-scrum}
\lessonslearned{
	\item Motivation for agile development
	\item Agile Manifesto (four values, twelve principles)
	\item Scrum: roles, terms, discussion
	\item Next: Are we developing the right product?
}{
	\item[] \sommerville\mychapter{3} Agile Software Development
}{
	\centering Want to learn how software is developed in practice? Lego-Scrum will take place on December 11, 2025 from 1--5pm!
	
	~
	
	\only<3|handout:0>{\pic[width=\linewidth,trim={0\width} {.1111\height} {0\width} {0\height},clip]{legoscrum/legoscrum2025-06-25result}}% translate from 4:3 into 3:2 format
	\only<4|handout:0>{\pic[width=\linewidth]{legoscrum/legoscrum2025-06-25table}}%
	\only<5->{\pic[width=\linewidth,trim={0\width} {0\height} {.2\width} {.2\height},clip]{legoscrum/legoscrum2025-06-25meeting}}%
}

%\faq{
	%	\item
	%}{
	%	\item
	%}{
	%	\item
	%}

\input{template/footer}
